### 2021-09-30 latex/appendix.tex
make lyx table only

\appendix

\section{Notes}

\subsection{To be investigated}

\begin{enumerate}

	\item [WIP]
		Sampling also matters in adversarial training for deep metric learning.

		[source hardness and destination hardness] different source sampling
		strategy and different destination sampling strategy. attack
		destination. DML loss not that sensitive to sampling? or generic
		advtrain method for DML? | still investigating

	\item [WIP] 7-step PGD and better aligned comparison: compare when
		these methods reach similar retrieval performance.

	\item Adversarial training (does not allow any distance moving) may
		encourage collapse (a trivial solution to the requirement). What if we
		loosen the requirement and convert it into a new constraint? For
		instance, \[
			L_{loosen}(a,p,n)=L_{base}(a,p,n)+\lambda_{1}L(a,\tilde{a},p)+\lambda_{2}L_{triplet}(a,\tilde{a},n)
		\] where the $L_{base}$ can be either $L_{triplet}$, $L_{rest}$, or
		$L_{act}$. Question: should we do this for every sample? i.e., \[
			L_{loosen}(a,p,n)=L_{base}(a,p,n)+\sum_{x\in\{a,p,n\}}L(x,\tilde{x},n_{x})
		\]

	\item ?  what about other types of metric learining loss functions? |
		investigating (we first make the most classical method work)

	\item ?  effectiveness of AMD Semi is to be verified. | as effective
		as ACT, reaches slightly higher ERS. But performs worse than ACT on
		some attacks, while performs better one some other attacks. | needs to
		be interpreted.

	\item faster training (algorithm tweak) FAT/other | postpone

	\item ? can we learn a good destination distribution? | too
		complicated? not necessary currently.

	\item 1. APGD as the optimization algorithm? | postpone

	\item 1. ? mixing amdsemi and act for adversarial training. They are good
		at different aspects.

	\item verify model collapse

	\item [\cmark] what if we use stopat for RAMD? | similar effect to amdsemi.
		misguiding gradient does not significantly affect AMD defense. So it
		is not quite necessary.

	\item [\cmark] \checkmark dynamically changing schedule for amdsemi? |
		looks good | the sqrt scheme works the best. (gradually increasing the
		hardness makes sense)

	\item [\cmark] implement FAT for faster exp iteration | works on toy
		dataset. too prone to collapse. if we try to avoid model collapse, the
		adversarial robustness will decrease significantly.

	\item [\cmark] (codecheck) does amd really honour the pgditer parameter? |
		direct adoptation of madry defense is really prone to model collapse.

\end{enumerate}

\subsection{Discarded or Rejected}

\begin{enumerate}

	\item \xmark
		MSE attack? | TMA is already close enough to it.

\end{enumerate}