### 2021-10-05 latex/sort-hardness.tex
tables

\begin{table}
\resizebox{\columnwidth}{!}{
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
Dataset & Random (r) & Semihard (m) & Softhard (s) & Distance (d) & Hardest (h)\tabularnewline
\hline 
\hline 
MNIST & -0.24 & -0.12 & -0.05 & 0.02 & 0.02\tabularnewline
\hline 
Fashion & -0.32 & -0.12 & -0.04 & 0.03 & 0.03\tabularnewline
\hline 
CUB & -0.15 & -0.13 & -0.10 & 0.04 & 0.05\tabularnewline
\hline 
CARS &  &  &  &  & \tabularnewline
\hline 
SOP &  &  &  &  & \tabularnewline
\hline 
\end{tabular}
}
\caption{(freeze) Sorting Hardness using average loss of first several (we
set 5) iterations.}
\end{table}