### 2021-10-26 tab-desth.tex
tex

\begin{table}
\resizebox{\columnwidth}{!}{
\setlength{\tabcolsep}{0.36em}
\begin{tabular}{c|cc|cc|cc|cc|cc}
	\toprule

	\multirow{2}{*}{\backslashbox{$H_\mathsf{S}$}{$H_\mathsf{D}$}} & \multicolumn{2}{c|}{Random} & \multicolumn{2}{c|}{Semihard} & \multicolumn{2}{c|}{Softhard} & \multicolumn{2}{c|}{Distance} & \multicolumn{2}{c}{Hardest}\tabularnewline
\cline{2-11} \cline{3-11} \cline{4-11} \cline{5-11} \cline{6-11} \cline{7-11} \cline{8-11} \cline{9-11} \cline{10-11} \cline{11-11} 
& R@1 & ERS & R@1 & ERS & R@1 & ERS & R@1 & ERS & R@1 & ERS\tabularnewline

\midrule

Random & \cellcolor{black!10}53.9 & \cellcolor{black!10}3.8 & 27.0 & 35.1 & \multicolumn{2}{c|}{Collapse} & \multicolumn{2}{c|}{Collapse} & \multicolumn{2}{c}{Collapse}\tabularnewline
\hline
Semihard & 43.9 & 5.4 & \cellcolor{black!10}44.0 & \cellcolor{black!10}5.0 & \multicolumn{2}{c|}{Collapse} & \multicolumn{2}{c|}{Collapse} & \multicolumn{2}{c}{Collapse}\tabularnewline
\hline
Softhard & 48.3 & 13.7 & 38.4 & 29.6 & \cellcolor{black!10}55.7 & \cellcolor{black!10}6.2 & \multicolumn{2}{c|}{Collapse} & \multicolumn{2}{c}{Collapse}\tabularnewline
\hline
Distance & 52.7 & 4.8 & 50.7 & 4.8 & \multicolumn{2}{c|}{Collapse} & \cellcolor{black!10}51.4 & \cellcolor{black!10}4.9 & 54.7 & 5.4\tabularnewline
\hline
Hardest & 51.0 & 4.7 & 52.2 & 4.8 & \multicolumn{2}{c|}{Collapse} & 52.6 & 5.1 & \cellcolor{black!10}48.9 & \cellcolor{black!10}5.0\tabularnewline

	\bottomrule
\end{tabular}
}

	\caption{Combinations of Source \& Destination Hardness.
	%
	Evaluated on CUB Dataset with RN18 model.
	%
	The last-epoch performance is reported instead of the peak performance.
	%
	Models on the diagonal are regularly (instead of adversarially) trained.
	%
}
	\label{tab:desth}

\end{table}