### 2021-11-06 tab-hsort.tex
p7

\begin{table}
\resizebox{\columnwidth}{!}{
	\setlength{\tabcolsep}{0.8em}
\begin{tabular}{c|ccccc}
\toprule
\rowcolor{black!7} Statistics & Random & Semihard & Softhard & Distance & Hardest\tabularnewline
\midrule
$E[H]$ & $-0.164$ & $-0.126$ & $-0.085$ & $0.043$ & $0.044$\tabularnewline
$\text{Var}[H]$ & $0.00035$ & $0.00013$ & $0.00122$ & $0.00021$ & $0.00021$\tabularnewline
\bottomrule
\end{tabular}
}
	\vspace{-0.5em}
\caption{Mean \& Variance of Hardness w/ Various Triplet Samplers.
%
Calculated as the average statistics over $1000$ mini-batches from CUB dataset
with an imagenet-initialized RN18 model.
%
}

	\label{tab:hsort}
\end{table}